Trends and Challenges in Toronto’s Homeless Population (2018-2024)

### 2024-09-27 paper/paper.qmd
Final commit

Analyzing age, gender and chronic homelessness helps to understand Toronto's homeless population. Homelessness is concentrated among working-age adults, particularly those aged 35-44 and 25-34, likely due to challenges with employment or housing stability. The sharp decline in the under-16 group during the pandemic may reflect government support for youth, but the subsequent increase underscores the need for sustained policies. In terms of gender, men remain the majority, with only 1.4% of people being transgender or non-binary. This suggests that shelters should design facilities to meet the different needs of men and women, while also accommodating transgender or non-binary people. The increase in the number of chronically homeless people highlights the inadequacy of current interventions to help people escape homelessness in the long term and signals a worsening situation without further intervention.

>>>>

Analyzing age, gender and chronic homelessness provides insight into Toronto's homeless population. Homelessness is concentrated among working-age adults, particularly the 35-44 and 25-34-year-olds, which may be due to challenges they face with employment or housing stability. During the pandemic, there was a sharp fall in homelessness among the under-16s group, which may reflect government support for young people during the pandemic. However, subsequent increases highlight the need for sustained policies.

In terms of gender, males remain in the majority (61.8\%) and only 1.4\% are transgender or non-binary, with little fluctuation in gender breakdown. In the long term, shelters should build facilities that are targeted to the male/female ratio. At the same time, the needs of transgender and non-binary people cannot be ignored, even if they make up only a small percentage of the population.

The increase in the number of chronically homeless people is largely due to the continued rise in rents and house prices, which are far beyond the ability of many low-income residents to pay, leaving them without the means to maintain stable accommodation [@whitzman2023human]. The rise in the proportion of people who are chronically homeless means that current interventions are not sufficient to help people out of homelessness and that the underlying issues need to be further addressed.
